### 2024-06-12 products/manuscript/Manuscript.qmd
Part 1 Final Project
Updated manuscript.qmd for part 1 of final project submission

---
title: "Parameters associated with Injuries in the NFL"
subtitle: ""
author: Brody Johnson
date: "`r Sys.Date()`"
format:
  docx:
    toc: false
    number-sections: true
    highlight-style: github
bibliography: ../../assets/dataanalysis-references.bib
csl: ../../assets/apa.csl
---

The structure below is one possible setup for a data analysis project (including the course project). For a manuscript, adjust as needed. 
You don't need to have exactly these sections, but the content covering those sections should be addressed.

This uses MS Word as output format. [See here](https://quarto.org/docs/output-formats/ms-word.html) for more information. You can switch to other formats, like html or pdf. See [the Quarto documentation](https://quarto.org/) for other formats.

```{r, echo=FALSE, message=FALSE}
# load a few R packages
library(here)
library(knitr)
```

# Summary/Abstract

The goal of this project will be to find variables related to injuries observed in the NFL. That is, find the common play types, game situation (time remaining, down, yards to go, etc.), and injured player metrics (position, height, body weight, etc.) that has a correlation to causing the resulting injury.

{{< pagebreak >}}

# Introduction

## General Background Information

For years, the NFL has been making changes in the game in an attempt to prevent player injuries. These changes range from rule modifications to investment in advancements to the players pads and helmets. Hopefully, the findings from this project can help inform players, coaches, and who knows, the NFL about the metrics for a higher risk in injury.

## Description of data and data source

Three data sets have been identified for this project. The primary data set ("Injury Player Data" obtained from a Github repository contain 1,586 obervations) consists of every play ran in the NFL containing an injury during the 2019-2020 season. Some of the things contained in this data include the teams playing, the week in the season, play description, injury area, and injured player metrics. Two other data sets were obtained from Kaggle. The first is a "cumulative play data" containing information about every play ran in the 2019-2020 NFL season (42,186 observations). Information such as the teams playing, play description, and time in the game when it was ran. The last data set is the "player demographic data" containing information about all NFL players in the 2015-2019 seasons (11,145 observations). This data set has player metrics such as height, weight, date of birth, etc.

## Questions/Hypotheses to be addressed
_State the research questions you plan to answer with this analysis._
My hope is to answer the following question: What metrics, if any, are available that can help predict whether an injury may occur? In turn, this has the potential to answer more questions. For example: 1. For coaches, are there specific play types that lead to more injuries? 2. For players, is there a goal weight for certain positions that can reduce the likelihood of injury?

To cite other work (important everywhere, but likely happens first in introduction), make sure your references are in the bibtex file specified in the YAML header above and have the right bibtex key. Then you can include like this:

Examples of reproducible research projects can for instance be found in [@mckay2020; @mckay2020a].

{{< pagebreak >}}

# Methods

_Describe your methods. That should describe the data, the cleaning processes, and the analysis approaches. You might want to provide a shorter description here and all the details in the supplement._

In both the "Injury Player Data" set and "cumulative play data" set, the play descriptions will need to be boiled down to simplified play types (pass, run, fumble, etc.). Next, a lot of what looks like unnecessary plays are included in the "cumulative play data" set such as end of quarters and games so these observations can be filtered out. Lastly, empty columns were noticed in the "cumulative play data" that can be removed.

## Schematic of workflow

Sometimes you might want to show a schematic diagram/figure that was not created with code (if you can do it with code, do it). @fig-schematic is an example of some - completely random/unrelated - schematic that was generated with Biorender.
We store those figures in the `assets` folder.

```{r}
#| label: fig-schematic
#| fig-cap: "A figure that is manually generated and shows some overview/schematic. This has nothing to do with the data, it's just a random one from one of our projects I found and placed here."
#| echo: FALSE
knitr::include_graphics(here("assets","antigen-recognition.png"))
```

## Data aquisition
_As applicable, explain where and how you got the data. If you directly import the data from an online source, you can combine this section with the next._

## Data import and cleaning
_Write code that reads in the file and cleans it so it's ready for analysis. Since this will be fairly long code for most datasets, it might be a good idea to have it in one or several R scripts. If that is the case, explain here briefly what kind of cleaning/processing you do, and provide more details and well documented code somewhere (e.g. as supplement in a paper). All materials, including files that contain code, should be commented well so everyone can follow along._

```{r}
#Loading the Data Sets
#install.packages("RCurl")
#library(RCurl)

# Injury Player Data
#injuries <- read.csv(text = getURL("https://raw.githubusercontent.com/btj5z2/BrodyJohnson-P2-portfolio/main/position_injury_player_2019_2020_update.csv")) #also works but not needed
injuries=read.csv("../../data/raw-data/position_injury_player_2019_2020_update.csv")

# Cumulative Play data
#pbp <- read.csv(text = getURL("https://raw.githubusercontent.com/btj5z2/BrodyJohnson-P2-portfolio/main/pbp-2019.csv")) #short for play-by-play #also works but not needed
pbp = read.csv("../../data/raw-data/pbp-2019.csv")

#Player Demographic data
#players <- read.csv(text = getURL("https://raw.githubusercontent.com/btj5z2/BrodyJohnson-P2-portfolio/main/player_position_Rmd1.csv"))
players = read.csv("../../data/raw-data/players (1).csv")

```

## Statistical analysis
_Explain anything related to your statistical analyses._
The big picture look at the analysis will be to compare the frequency of parameters in the injury data set to the same parameters in the two supporting data sets. For example, we may find in the injury data set that most injuries occur during "short pass" plays. However, if we were to find the "short pass" play is the most common play type ran in the NFL, this becomes less significant of a finding.

{{< pagebreak >}}

# Results

## Exploratory/Descriptive analysis

_Use a combination of text/tables/figures to explore and describe your data. Show the most important descriptive results here. Additional ones should go in the supplement. Even more can be in the R and Quarto files that are part of your project._

### Structures of each data set

```{r}
#Structure of Injury Data Set
str(injuries)
```

```{r}
#Structure of Play-by-Play Data
str(pbp)
```

```{r}
#Structure of All Players Data Set
str(players)
```

### Exploratory Analysis of parameters only in the *Injury* Data Set

```{r}
hist(injuries$week.x, breaks=24,main="Injuries by Week in Regular Season", xlim=c(1,18),xlab="Week",labels =T)
```

### Exploratory Analysis of parameters in the *Injury* Data Set and *Play-by-play* Data Set

```{r}
par(mfrow=c(1,2))
hist(injuries$qtr, breaks=6,main="Injuries by Quarter", xlim=c(0,5),xlab="Quarter",labels =T)
hist(pbp$Quarter, breaks=6,main="Total Plays by Quarter", xlim=c(0,5), xlab="Quarter",labels =T)

```

```{r}
par(mfrow=c(1,2))
hist(injuries$down, breaks=6,main="Injuries by Down", xlim=c(0,5),xlab="Quarter",labels =T)
hist(pbp$Down, breaks=6,main="Total Plays by Down", xlim=c(0,5), xlab="Quarter",labels =T)

```

### Exploratory Analysis of parameters in the *Injury* Data Set and *Cumulative Player* Data Set

```{r}
par(mfrow=c(1,2))
hist(injuries$height, breaks=10,main="Injuries by Height", xlab="Heiht",labels =T)
hist(as.numeric(players$height), breaks=10,main="Total Players Height", xlab="Height",labels =T) #Nedd to look into why not all numeric

```

```{r}
par(mfrow=c(1,2))
hist(injuries$weight, breaks=10,main="Injuries by Weight", xlab="Weight",labels =T)
hist(players$weight, breaks=10,main="Total Players Weight", xlab="Weight",labels =T)

```

@tbl-summarytable shows a summary of the data.

Note the loading of the data providing a **relative** path using the `../../` notation. (Two dots means a folder up). You never want to specify an **absolute** path like `C:\ahandel\myproject\results\` because if you share this with someone, it won't work for them since they don't have that path. You can also use the `here` R package to create paths. See examples of that below. I generally recommend the `here` package.

```{r}
#| label: tbl-summarytable
#| tbl-cap: "Data summary table."
#| echo: FALSE
resulttable=readRDS("../../results/tables/summarytable.rds")
knitr::kable(resulttable)
```

## Basic statistical analysis

_To get some further insight into your data, if reasonable you could compute simple statistics (e.g. simple models with 1 predictor) to look for associations between your outcome(s) and each individual predictor variable. Though note that unless you pre-specified the outcome and main exposure, any "p<0.05 means statistical significance" interpretation is not valid._

@fig-result shows a scatterplot figure produced by one of the R scripts.

```{r}
#| label: fig-result
#| fig-cap: "Height and weight stratified by gender."
#| echo: FALSE
knitr::include_graphics(here("results","figures","height-weight-stratified.png"))
```

## Full analysis

_Use one or several suitable statistical/machine learning methods to analyze your data and to produce meaningful figures, tables, etc. This might again be code that is best placed in one or several separate R scripts that need to be well documented. You want the code to produce figures and data ready for display as tables, and save those. Then you load them here._

Example @tbl-resulttable2 shows a summary of a linear model fit.

```{r}
#| label: tbl-resulttable2
#| tbl-cap: "Linear model fit table."
#| echo: FALSE
resulttable2 = readRDS(here("results","tables","resulttable2.rds"))
knitr::kable(resulttable2)
```

{{< pagebreak >}}

# Discussion

## Summary and Interpretation
_Summarize what you did, what you found and what it means._

## Strengths and Limitations
_Discuss what you perceive as strengths and limitations of your analysis._

## Conclusions
_What are the main take-home messages?_

_Include citations in your Rmd file using bibtex, the list of references will automatically be placed at the end_

This paper [@leek2015] discusses types of analyses.

These papers [@mckay2020; @mckay2020a] are good examples of papers published using a fully reproducible setup similar to the one shown in this template.

Note that this cited reference will show up at the end of the document, the reference formatting is determined by the CSL file specified in the YAML header. Many more style files for almost any journal [are available](https://www.zotero.org/styles). You also specify the location of your bibtex reference file in the YAML. You can call your reference file anything you like.

{{< pagebreak >}}

# References

### 2024-06-17 products/manuscript/Manuscript.qmd
Cleaned data sets

@@
 Three data sets have been identified for this project.
-The primary data set ("Injury Player Data" obtained from a Github repository contain 1,586 obervations) consists of every play ran in the NFL containing an injury during the 2019-2020 season.
+The primary data set ("Injury Player Data" obtained from a Github repository contain 1,586 observations) consists of every play ran in the NFL containing an injury during the 2019-2020 season.
 Some of the things contained in this data include the teams playing, the week in the season, play description, injury area, and injured player metrics.
@@
 ## Questions/Hypotheses to be addressed
-_State the research questions you plan to answer with this analysis._
+
 My hope is to answer the following question: What metrics, if any, are available that can help predict whether an injury may occur?

### 2024-06-17 code/processing-code/processingcode.R

@@ -22,100 +22,216 @@ library(here) #to set paths
 ## ---- loaddata --------
 #path to data
 #note the use of the here() package and not absolute paths
-data_location <- here::here("data","raw-data","exampledata.xlsx")
+#data_location <- here::here("data","raw-data","exampledata.xlsx")
 #load data. 
 #note that for functions that come from specific packages (instead of base R)
 # I often specify both package and function like so
 #package::function() that's not required one could just call the function
 #specifying the package makes it clearer where the function "lives",
 #but it adds typing. You can do it either way.
-rawdata <- readxl::read_excel(data_location)
+#rawdata <- readxl::read_excel(data_location)
 # We might also want to load the codebook to look at it
-codebook <- readxl::read_excel(data_location, sheet ="Codebook")
+#codebook <- readxl::read_excel(data_location, sheet ="Codebook")
+#Loading the Data Sets
+#install.packages("RCurl")
+#library(RCurl)
 
+# Injury Player Data
+#injuries <- read.csv(text = getURL("https://raw.githubusercontent.com/btj5z2/BrodyJohnson-P2-portfolio/main/position_injury_player_2019_2020_update.csv")) #also works but not needed
+injuries=read.csv("data/raw-data/position_injury_player_2019_2020_update.csv")
+
+# Cumulative Play data
+#pbp <- read.csv(text = getURL("https://raw.githubusercontent.com/btj5z2/BrodyJohnson-P2-portfolio/main/pbp-2019.csv")) #short for play-by-play #also works but not needed
+pbp = read.csv("data/raw-data/pbp-2019.csv")
+
+#Player Demographic data
+#players <- read.csv(text = getURL("https://raw.githubusercontent.com/btj5z2/BrodyJohnson-P2-portfolio/main/player_position_Rmd1.csv"))
+players = read.csv("data/raw-data/players (1).csv")
 
 ## ---- exploredata --------
 #take a look at the data
-dplyr::glimpse(rawdata)
+dplyr::glimpse(injuries)
+  #Summarize "description" into "play_type"
+  #Summarize "injury_area" as upper vs lower body into new parameter
+  #Create new field calculating age from birthdate (mm/dd/yyyy)
+dplyr::glimpse(pbp) 
[…]

### 2024-06-22 products/manuscript/Manuscript.qmd
Submission part 3 changes

@@
 library(knitr)
+library(dplyr)
+library(tidyverse)
 ```
@@
 Three data sets have been identified for this project.
-The primary data set ("Injury Player Data" obtained from a Github repository contain 1,586 observations) consists of every play ran in the NFL containing an injury during the 2019-2020 season.
+The primary data set ("Injury Player Data" obtained from a Github repository contains 1,586 observations) consists of every play ran in the NFL containing an injury during the 2019-2020 season.
 Some of the things contained in this data include the teams playing, the week in the season, play description, injury area, and injured player metrics.
@@
 Information such as the teams playing, play description, and time in the game when it was ran.
-The last data set is the "player demographic data" containing information about all NFL players in the 2015-2019 seasons (11,145 observations).
+The last data set is the "player demographic data" containing information about all NFL players in the 2016-2019 seasons (11,145 observations).
 This data set has player metrics such as height, weight, date of birth, etc.
@@
 
-To cite other work (important everywhere, but likely happens first in introduction), make sure your references are in the bibtex file specified in the YAML header above and have the right bibtex key.
+*****To cite other work (important everywhere, but likely happens first in introduction), make sure your references are in the bibtex file specified in the YAML header above and have the right bibtex key.
 Then you can include like this:
 
-Examples of reproducible research projects can for instance be found in [@mckay2020; @mckay2020a].
+Examples of reproducible research projects can for instance be found in [@mckay2020; @mckay2020a].******
 
@@
 
-_Describe your methods. That should describe the data, the cleaning processes, and the analysis approaches. You might want to provide a shorter description here and all the details in the supplement._
+The primary idea behind the analysis for this project is to look at various parameters when an injury occurred and compare them to the rest of the NFL.
+If the parameters are related to the play itself, the paramter in the injruy data set will be compared to cumulative play data set.
+Likewise, if the paramter is related to the injured player themselves, this paramter will be copmared to the cumulative player data set.
+If there is a statistical difference between paramters, we can likely draw some conclusion about a parameter causing injuries.
 
-In both the "Injury Player Data" set and "cumulative play data" set, the play descriptions will need to be boiled down to simplified play types (pass, run, fumble, etc.).
-Next, a lot of what looks like unnecessary plays are included in the "cumulative play data" set such as end of quarters and games so these observations can be filtered out.
-Lastly, empty columns were noticed in the "cumulative play data" that can be removed.
+With this being said, it is imperative to ensure parameters from both data sets are in the same format in order to compare them.
+Other data cleaning items will include vetting the data sets for unexpected/missing values and converting variables to another variable that easier to use.
+For example, the play descriptions in the injury data set and cumulative play data set are very extensive in what happened during the play.
+Therefore, this will be boiled down to be either a run play, short pass, etc.
 
@@
 ## Data import and cleaning
-_Write code that reads in the file and cleans it so it's ready for analysis.
-Since this will be fairly long code for most datasets, it might be a good idea to have it in one or several R scripts.
-If that is the case, explain here briefly what kind of cleaning/processing you do, and provide more details and well documented code somewhere (e.g.
-as supplement in a paper).
-All materials, including files that contain code, should be commented well so everyone can follow along._
+In both the "Injury Player Data" set and "cumulative play data" set, the play descriptions will need to be boiled down to simplified play types (pass, run, fumble, etc.).
+Next for the "Injury Player Data" set, the "injury_area" parameter will be summarized as upper and lower body injuries.
+This is due to the overwhelming number of injuries being classified as one or the other already so this will classify the remaining injuries as such.
+For the injury data set and the "player demographic data", player ages will need to be calculated from their birth-dates.
+
+As for the "cumulative player data" set, empty columns will be removed and 0's in the down column will be converted to "NA" values which indicate kickoffs & PATs.
+Lastly, a lot of unnecessary plays are included in the "cumulative play data" set such as end of quarter/games, 2-minute warnings, and timeouts so these observations can be filtered out.
+
+In the "player demographic data" set, observations for players height were recorded in both inches and feet-inches so this was cleaned to only contain data in inches.
+Lastly, the players position was converted to a factor for analysis.
 
 ```{r}
-#Loading the Data Sets
-#install.packages("RCurl")
-#library(RCurl)
+#Loading the processed Data Sets
+#Raw data files were cleaned in the processingcode.R (../../code/processing-code/) then saved as .rds file in the processed-data folder
 
 # Injury Player Data
-#injuries <- read.csv(text = getURL("https://raw.githubusercontent.com/btj5z2/BrodyJohnson-P2-portfolio/main/position_injury_player_2019_2020_update.csv")) #also works but not needed
-injuries=read.csv("../../data/raw-data/position_injury_player_2019_2020_update.csv")
+injuries=readRDS("../../data/processed-data/injuriesprocesseddata.rds")
 
 # Cumulative Play data
-#pbp <- read.csv(text = getURL("https://raw.githubusercontent.com/btj5z2/BrodyJohnson-P2-portfolio/main/pbp-2019.csv")) #short for play-by-play #also works but not needed
-pbp = read.csv("../../data/raw-data/pbp-2019.csv")
+pbp=readRDS("../../data/processed-data/pbpprocesseddata.rds")
 
 #Player Demographic data
-#players <- read.csv(text = getURL("https://raw.githubusercontent.com/btj5z2/BrodyJohnson-P2-portfolio/main/player_position_Rmd1.csv"))
-players = read.csv("../../data/raw-data/players (1).csv")
-
+players=readRDS("../../data/processed-data/playersprocesseddata.rds") 
 ```
@@
 ## Statistical analysis
-_Explain anything related to your statistical analyses._
-The big picture look at the analysis will be to compare the frequency of parameters in the injury data set to the same parameters in the two supporting data sets.
-For example, we may find in the injury data set that most injuries occur during "short pass" plays.
-However, if we were to find the "short pass" play is the most common play type ran in the NFL, this becomes less significant of a finding.
+First, I will select variables that show a potential difference between the injury data set and cumulative  data sets through exploratory analysis.
+From these variables, I will run  inferential tests for the numeric variables and a Chi-square test for the categorical variables.
 
@@
 
-_Use a combination of text/tables/figures to explore and describe your data.
-Show the most important descriptive results here.
-Additional ones should go in the supplement.
-Even more can be in the R and Quarto files that are part of your project._
-
 ### Structures of each data set
@@
 #Structure of Injury Data Set
-str(injuries)
+glimpse(injuries)
 ```
@@
 #Structure of Play-by-Play Data
-str(pbp)
+glimpse(pbp)
 ```
@@
 #Structure of All Players Data Set
-str(players)
+glimpse(players)
 ```
@@
 
+Looking at the number of injuries each week in the regular season, there's a sharp drop after the first week, steadily climbs nearing the end of the season, and another drop at the end of the season.
+Intuitively, this makes sense in that players may not be physically readily for the intensity of a regular-season game despite training camps and preseason games.
+After that, the slight rise in the second half of the season could be due to exhaustion through a long season and/or teams fighting more for playoff spots.
+Lastly, the steep drop in the last week is likely due to teams knowing whether or not they have made playoffs and therefore, do not want to incur an injury to go into either the post-season or off-season.
 ```{r}
-hist(injuries$week.x, breaks=24,main="Injuries by Week in Regular Season", xlim=c(1,18),xlab="Week",labels =T)
+hist(injuries$week.x, breaks=21,main="Injuries by Week in Regular Season", xlim=c(1,17),ylim=c(0,200),xlab="Week",labels =T)
+```
+
+Below depicts the number of injuries for each reported injury area. The top results were simply reported as upper or lower body.
+
+```{r}
+injarea_freq=injuries %>%  group_by(injury.area) %>% summarize(count=n()) 
+ggplot(data = injarea_freq, aes(x = reorder(injury.area, -count), y = count))+
+  geom_col() +
+  labs(x = "Injury Area", y="# of Injuries") + 
+  theme(axis.text.x = element_text(angle=45, hjust = 1))
+
+```
+
+Below is a plot of the number of injuries by injury area but the injury areas were summarized to either be upper/lower body or head injuries.
+I kept head injuries separate due to the interest around concussions in football and see how they stack up against all other injuries.
+
+```{r}
+injareanew_freq=injuries %>%  group_by(injury.area.new) %>% summarize(count=n()) 
+ggplot(data = injareanew_freq, aes(x = reorder(injury.area.new, -count), y = count))+
+  geom_col() +
+  labs(x = "Injury Area", y="# of Injuries") + 
+  theme(axis.text.x = element_text(angle=45, hjust = 1))
+```
+
+Out of curiosity, I plotted below the frequency of injuries for each team during the regular season.
+I was not expecting much but it is interesting to see about a triple in number of injuries from the lowest team (LA Rams) and the highest (LA Chargers).
+
+```{r}
+team_freq=injuries %>%  filter(week.x<18) %>% group_by(team) %>% summarize(count=n()) #filtered to only contain regular season games (weeks 1-17)
+ggplot(data = team_freq, aes(x = reorder(team, -count), y = count))+
+  geom_col() +
+  labs(x = "Team", y="# of Injuries") + 
+  theme(axis.text.x = element_text(angle=45, hjust = 1))
 ```
@@
 
+The number of plays seem evenly distributed among the halves (slight increase in 2nd and 4th quarters likely due to running last minute plays).
+However, the number of injuries consistently increase throughout the game which also makes sense as players fatigue.
+
 ```{r}
-par(mfrow=c(1,2))
-hist(injuries$qtr, breaks=6,main="Injuries by Quarter", xlim=c(0,5),xlab="Quarter",labels =T)
-hist(pbp$Quarter, breaks=6,main="Total Plays by Quarter", xlim=c(0,5), xlab="Quarter",labels =T)
+hist(injuries$qtr,main="Injuries by Quarter", ylim=c(0,550), xlim=c(1,5),xlab="Quarter",labels =T)
+
+```
+```{r}
+hist(pbp$Quarter, breaks=6,main="Total Plays by Quarter", ylim=c(0,11200), xlab="Quarter",labels =T)
+
+```
+
+There appears to be a very similar trend among the number of plays per down and the number of injuries ocurring during each down. That is, they both decrease at similar rates.
+
+```{r}
+hist(injuries$down, breaks=5,main="Injuries by Down", ylim=c(0,700),xlim=c(1,4),xlab="Down",labels =T)
+```
+
+```{r}
+hist(pbp$Down, breaks=6,main="Total Plays by Down", ylim=c(0,14900),xlim=c(1,4), xlab="Down",labels =T)
+```
+
+Below are two histograms depicting first the number of injuries and then second the total # of plays by different play types.
+After a glance, there does not appear to be a significant difference in number of injuries and total # of plays ran by play type.
+
+```{r}
+play_type_freq=injuries %>%  group_by(play_type) %>% summarize(count=n()) 
+ggplot(data = play_type_freq, aes(x = reorder(play_type, -count), y = count))+
+  geom_col() +
+  labs(x = "Play Type", y="# of Injuries") + 
+  theme(axis.text.x = element_text(angle=45, hjust = 1))
 
@@
 ```{r}
-par(mfrow=c(1,2))
-hist(injuries$down, breaks=6,main="Injuries by Down", xlim=c(0,5),xlab="Quarter",labels =T)
-hist(pbp$Down, breaks=6,main="Total Plays by Down", xlim=c(0,5), xlab="Quarter",labels =T)
+play_type_freq2=pbp %>%  group_by(play_type) %>% summarize(count=n()) 
+ggplot(data = play_type_freq2, aes(x = reorder(play_type, -count), y = count))+
+  geom_col() +
+  labs(x = "Play Type", y="Total # of Plays") + 
+  theme(axis.text.x = element_text(angle=45, hjust = 1))
 
@@
 
+Below are the distributions of players height first of the injured players and then from the cumulative player data set.
+Both histograms appear to be normally distributed.
+One interesting difference is the injured players appear to have more players just below and above the center of the histogram.
+That is, it would almost suggest players specifically under and above the mean height have a higher inclination for injury.
+
 ```{r}
-par(mfrow=c(1,2))
-hist(injuries$height, breaks=10,main="Injuries by Height", xlab="Heiht",labels =T)
-hist(as.numeric(players$height), breaks=10,main="Total Players Height", xlab="Height",labels =T) #Nedd to look into why not all numeric
+hist(injuries$height, breaks=10,main="Injuries by Height", ylim=c(0,250), xlab="Height",labels =T)
+
+```
+```{r}
+hist(players$height, breaks=10,main="Total Players Height", ylim=c(0,250), xlab="Height",labels =T)
+
+```
+
+Below are the histograms depicting the players height. They are both skewed to the right with one interesting finding of a peak of injured players in the ~300lb range.
+
+```{r}
+hist(injuries$weight, breaks=10,main="Injuries by Weight", xlim=c(150,400),ylim=c(0,400), xlab="Weight",labels =T)
+```
+
+```{r}
+hist(players$weight, breaks=10,main="Total Players Weight", xlim=c(150,400),ylim=c(0,400),  xlab="Weight",labels =T) 
+```
+
+Both distributions of players age are below and have a similar skewness to the right.
+
+```{r}
+hist(injuries$age, breaks=10,main="Injuries by Age", xlim=c(20,45),ylim=c(0,425), xlab="Age",labels =T)
+```
+
+```{r}
+hist(players$age, breaks=10,main="Total Players Age", xlim=c(20,45),ylim=c(0,425),  xlab="Age",labels =T) 
+```
+
+The primary finding from the player positions is that it appears the linemen (defensive and ofensive) are injured more often relative to how many there are in the league.
+
+```{r}
+position_freq=injuries %>%  group_by(position) %>% summarize(count=n()) 
+ggplot(data = position_freq, aes(x = reorder(position, -count), y = count))+
+  geom_col() +
+  labs(x = "Player Position", y="# of Injuries") + 
+  theme(axis.text.x = element_text(angle=45, hjust = 1))
 
@@
 ```{r}
-par(mfrow=c(1,2))
-hist(injuries$weight, breaks=10,main="Injuries by Weight", xlab="Weight",labels =T)
-hist(players$weight, breaks=10,main="Total Players Weight", xlab="Weight",labels =T)
+position_freq2=players %>%  group_by(position) %>% summarize(count=n()) 
+ggplot(data = position_freq2, aes(x = reorder(position, -count), y = count))+
+  geom_col() +
+  labs(x = "Player Position", y="Total # of Players") + 
+  theme(axis.text.x = element_text(angle=45, hjust = 1))
 
-```
-
-@tbl-summarytable shows a summary of the data.
-
-Note the loading of the data providing a **relative** path using the `../../` notation.
-(Two dots means a folder up).
-You never want to specify an **absolute** path like `C:\ahandel\myproject\results\` because if you share this with someone, it won't work for them since they don't have that path.
-You can also use the `here` R package to create paths.
-See examples of that below.
-I generally recommend the `here` package.
-
-```{r}
-#| label: tbl-summarytable
-#| tbl-cap: "Data summary table."
-#| echo: FALSE
-resulttable=readRDS("../../results/tables/summarytable.rds")
-knitr::kable(resulttable)
 ```
@@
 
-_To get some further insight into your data, if reasonable you could compute simple statistics (e.g.
-simple models with 1 predictor) to look for associations between your outcome(s) and each individual predictor variable.
-Though note that unless you pre-specified the outcome and main exposure, any "p<0.05 means statistical significance" interpretation is not valid._
-
-@fig-result shows a scatterplot figure produced by one of the R scripts.
+From the exploratory analysis, it appears there three numeric variables that may show a difference and one categorical.
+The three numeric variables are quarter, height, and weight while the single categorical variable is player position.
 
 ```{r}
-#| label: fig-result
-#| fig-cap: "Height and weight stratified by gender."
-#| echo: FALSE
-knitr::include_graphics(here("results","figures","height-weight-stratified.png"))
+#Inferential test looking at Quarter
+shapiro.test(injuries$qtr) #p-value<0.05. Reject Ho & don't assume normal.
+shapiro.test(sample(pbp$Quarter, size=5000, replace=F)) #p-value<0.05. Reject Ho & don't assume normal.
+#Will do Wilcoxon rank-sum test b/c of non-normal data
+
+wilcox.test(injuries$qtr, pbp$Quarter, exact=F) #p-value<0.05. Reject Ho and therefore, one group tends to have larger values than the other
+
+```
+
+```{r}
+#Inferential test looking at Height
+shapiro.test(injuries$height) #p-value<0.05. Reject Ho & don't assume normal.
+shapiro.test(players$height) #p-value<0.05. Reject Ho & don't assume normal.
+#Will do Wilcoxon rank-sum test b/c of non-normal data
+
+wilcox.test(injuries$height, players$height, exact=F) #p-value<0.05. Reject Ho and therefore, one group tends to have larger values than the other
+
+```
+
+```{r}
+#Inferential test looking at Weight
+shapiro.test(injuries$weight) #p-value<0.05. Reject Ho & don't assume normal.
+shapiro.test(players$weight) #p-value<0.05. Reject Ho & don't assume normal.
+#Will do Wilcoxon rank-sum test b/c of non-normal data
+
+wilcox.test(injuries$weight, players$weight, exact=F) #p-value<0.05. Reject Ho and therefore, one group tends to have larger values than the other
+
+```
+```{r}
+#Chi-square test looking at player position
+factor_levels = levels(injuries$position)
+chisq.test(sample(injuries$position, size=1303, replace=F), players$position) #p-value>0.05. Accept Ho and therefore, the two data sets have equal proportions among the player positions
+
 ```

### 2024-06-22 code/processing-code/processingcode.R

@@ -205,11 +205,27 @@ for (i in seq_along(pl1$height)) {
     pl1$height[i] = as.numeric(pl1$height[i])
   }
 }
-
+pl1$height = as.numeric(pl1$height) #Convert to numeric 
 
 ## ---- cleanplayersdata2 --------
-#Character parameters may also need to be converted to factor. 
-pl1$position <- as.factor(pl1$position)  
+#Create new field summarizing player positions to match the injury data set positions 
+pl1$position = as.character(pl1$position) #Had to convert to character b/c R wont add a new factor
+pl1$position = replace(pl1$position, pl1$position == 'SS', 'DB')
+pl1$position = replace(pl1$position, pl1$position == 'CB', 'DB')
+pl1$position = replace(pl1$position, pl1$position == 'FS', 'DB')
+pl1$position = replace(pl1$position, pl1$position == 'DE', 'DL')
+pl1$position = replace(pl1$position, pl1$position == 'DT', 'DL')
+pl1$position = replace(pl1$position, pl1$position == 'NT', 'DL')
+pl1$position = replace(pl1$position, pl1$position == 'OLB', 'LB')
+pl1$position = replace(pl1$position, pl1$position == 'ILB', 'LB')
+pl1$position = replace(pl1$position, pl1$position == 'MLB', 'LB')
+pl1$position = replace(pl1$position, pl1$position == 'HB', 'OL')
+pl1$position = replace(pl1$position, pl1$position == 'FB', 'OL')
+pl1$position = replace(pl1$position, pl1$position == 'P', 'SPEC')
+pl1$position = replace(pl1$position, pl1$position == 'S', 'SPEC')
+pl1$position = replace(pl1$position, pl1$position == 'K', 'SPEC')
+pl1$position = replace(pl1$position, pl1$position == 'LS', 'SPEC')
+pl1$position = as.factor(pl1$position) #convert back to factor 
 
 ## ---- cleanplayersdata3 --------
 #Create new field calculating age from birthdate (yyyy-mm-dd)
@@ -218,7 +234,7 @@ pl1$age = NA
 pl1$age = round(as.numeric(difftime(as.Date("2019-09-05"), pl1$birthDate, units = 'weeks'))/52, digits=1)
 
 
-## ---- savedata --------
+## ---- savedatas.character.srcref()## ---- savedata --------
 # all done, data is clean now. 
 # Let's assign at the end to some final variable
 # makes it easier to add steps above

### 2024-07-13 products/manuscript/Manuscript.qmd
Project submission #4

@@
 
-The structure below is one possible setup for a data analysis project (including the course project). For a manuscript, adjust as needed. 
-You don't need to have exactly these sections, but the content covering those sections should be addressed.
-
-This uses MS Word as output format.
-[See here](https://quarto.org/docs/output-formats/ms-word.html) for more information.
-You can switch to other formats, like html or pdf.
-See [the Quarto documentation](https://quarto.org/) for other formats.
-
 ```{r, echo=FALSE, message=FALSE}
@@
 library(tidyverse)
+library(ggplot2)
 ```
@@
 
-For years, the NFL has been making changes in the game in an attempt to prevent player injuries.
-These changes range from rule modifications to investment in advancements to the players pads and helmets.
+Over the past three decades, the NFL has come under increased scrutiny over the dangers of American football (@frontline-nfl-concussion).
+What initially began as concerns over the persistent brutality central to the sport, shifted into the more insidious issue of concussions and their role in the post-mortem diagnosis of Chronic Traumatic Encephalitic (CTE) (@mayoclinic-cte).
+The NFL has been making changes in the game in an attempt to prevent player injuries from rule modifications to investment in advancements to the players pads and helmets.
 Hopefully, the findings from this project can help inform players, coaches, and who knows, the NFL about the metrics for a higher risk in injury.
@@
 Three data sets have been identified for this project.
-The primary data set ("Injury Player Data" obtained from a Github repository contains 1,586 observations) consists of every play ran in the NFL containing an injury during the 2019-2020 season.
+The primary data set ("Injury Player Data" contains 1,586 observations) consists of every play ran in the NFL containing an injury during the 2019-2020 season.
 Some of the things contained in this data include the teams playing, the week in the season, play description, injury area, and injured player metrics.
-Two other data sets were obtained from Kaggle.
-The first is a "cumulative play data" containing information about every play ran in the 2019-2020 NFL season (42,186 observations).
+The next is a "cumulative play data" containing information about every play ran in the 2019-2020 NFL season (42,186 observations).
 Information such as the teams playing, play description, and time in the game when it was ran.
@@
 
-*****To cite other work (important everywhere, but likely happens first in introduction), make sure your references are in the bibtex file specified in the YAML header above and have the right bibtex key.
-Then you can include like this:
-
-Examples of reproducible research projects can for instance be found in [@mckay2020; @mckay2020a].******
-
 {{< pagebreak >}}
@@
 The primary idea behind the analysis for this project is to look at various parameters when an injury occurred and compare them to the rest of the NFL.
-If the parameters are related to the play itself, the paramter in the injruy data set will be compared to cumulative play data set.
-Likewise, if the paramter is related to the injured player themselves, this paramter will be copmared to the cumulative player data set.
-If there is a statistical difference between paramters, we can likely draw some conclusion about a parameter causing injuries.
+If the parameters are related to the play itself, the parameter in the injury data set will be compared to cumulative play data set.
+Likewise, if the parameter is related to the injured player themselves, this parameter will be compared to the cumulative player data set.
+If there is a statistical difference between parameters, we can likely draw some conclusion about a parameter causing injuries.
 
@@
 
-## Schematic of workflow
+## Data aquisition
+The "Injury Player Data" set was found from a Github repository.
+The repository belongs to a group of students who wanted to do a similar analysis looking at NFL injuries and various factors that could affect them.
 
-Sometimes you might want to show a schematic diagram/figure that was not created with code (if you can do it with code, do it).
-@fig-schematic is an example of some - completely random/unrelated - schematic that was generated with Biorender.
-We store those figures in the `assets` folder.
+https://github.com/sammieerne/NFL-Injury-Analysis/tree/main/Data
 
-```{r}
-#| label: fig-schematic
-#| fig-cap: "A figure that is manually generated and shows some overview/schematic. This has nothing to do with the data, it's just a random one from one of our projects I found and placed here."
-#| echo: FALSE
-knitr::include_graphics(here("assets","antigen-recognition.png"))
-```
+The "cumulative play data" set comes from a website called NFLsavant.com. This website is dedicated to providing NFL statistics to the public.
 
-## Data aquisition
-_As applicable, explain where and how you got the data. If you directly import the data from an online source, you can combine this section with the next._
+https://nflsavant.com/about.php
+
+Lastly, the "player demographic data" set was also taken from a Github repository. Similar to the other repository, this one belongs to a group who wanted to analyze injuries in the NFL.
+
+https://github.com/ericcrouse/NFL-Injury-Exploration/tree/main/data/raw
 
@@
 
-As for the "cumulative player data" set, empty columns will be removed and 0's in the down column will be converted to "NA" values which indicate kickoffs & PATs.
+As for the "cumulative play data" set, empty columns will be removed and 0's in the down column will be converted to "NA" values which indicate kickoffs & PATs.
 Lastly, a lot of unnecessary plays are included in the "cumulative play data" set such as end of quarter/games, 2-minute warnings, and timeouts so these observations can be filtered out.
@@
 ```{r}
-hist(injuries$qtr,main="Injuries by Quarter", ylim=c(0,550), xlim=c(1,5),xlab="Quarter",labels =T)
-
+ggplot(injuries, aes(x = qtr)) +
+  geom_histogram(binwidth = 1, fill = "lightgray", color = "black", alpha = 1) +
+  labs(title = "Injuries by Quarter", x = "Quarter", y = "Frequency")
 ```
 ```{r}
-hist(pbp$Quarter, breaks=6,main="Total Plays by Quarter", ylim=c(0,11200), xlab="Quarter",labels =T)
-
+ggplot(pbp, aes(x = Quarter)) +
+  geom_histogram(binwidth = 1, fill = "lightgray", color = "black", alpha = 1) +
+  labs(title = "Total Plays by Quarter", x = "Quarter", y = "Frequency")
 ```
@@
 ```{r}
-hist(injuries$down, breaks=5,main="Injuries by Down", ylim=c(0,700),xlim=c(1,4),xlab="Down",labels =T)
+ggplot(injuries, aes(x = down)) +
+  geom_histogram(binwidth = 1, fill = "lightgray", color = "black", alpha = 1) +
+  labs(title = "Injuries by Down", x = "Down", y = "Frequency")
 ```
@@
 ```{r}
-hist(pbp$Down, breaks=6,main="Total Plays by Down", ylim=c(0,14900),xlim=c(1,4), xlab="Down",labels =T)
+ggplot(pbp, aes(x = Down)) +
+  geom_histogram(binwidth = 1, fill = "lightgray", color = "black", alpha = 1) +
+  labs(title = "Total Plays by Down", x = "Down", y = "Frequency")
 ```
@@
 ```{r}
-hist(injuries$height, breaks=10,main="Injuries by Height", ylim=c(0,250), xlab="Height",labels =T)
-
+ggplot(injuries, aes(x = height)) +
+  geom_histogram(binwidth = 1, fill = "lightgray", color = "black", alpha = 1) +
+  labs(title = "Injuries by Height", x = "Height", y = "Frequency")
 ```
 ```{r}
-hist(players$height, breaks=10,main="Total Players Height", ylim=c(0,250), xlab="Height",labels =T)
-
+ggplot(players, aes(x = height)) +
+  geom_histogram(binwidth = 1, fill = "lightgray", color = "black", alpha = 1) +
+  labs(title = "Total Players by Height", x = "Height", y = "Frequency")
 ```
@@
 ```{r}
-hist(injuries$weight, breaks=10,main="Injuries by Weight", xlim=c(150,400),ylim=c(0,400), xlab="Weight",labels =T)
+ggplot(injuries, aes(x = weight)) +
+  geom_histogram(binwidth = 10, fill = "lightgray", color = "black", alpha = 1) +
+  labs(title = "Injuries by Weight", x = "Weight", y = "Frequency")
 ```
@@
 ```{r}
-hist(players$weight, breaks=10,main="Total Players Weight", xlim=c(150,400),ylim=c(0,400),  xlab="Weight",labels =T) 
+ggplot(players, aes(x = weight)) +
+  geom_histogram(binwidth = 10, fill = "lightgray", color = "black", alpha = 1) +
+  labs(title = "Total Players by Weight", x = "Weight", y = "Frequency")
 ```
@@
 ```{r}
-hist(injuries$age, breaks=10,main="Injuries by Age", xlim=c(20,45),ylim=c(0,425), xlab="Age",labels =T)
+ggplot(injuries, aes(x = age)) +
+  geom_histogram(binwidth = 1, fill = "lightgray", color = "black", alpha = 1) +
+  labs(title = "Injuries by Age", x = "Age", y = "Frequency")
 ```
@@
 ```{r}
-hist(players$age, breaks=10,main="Total Players Age", xlim=c(20,45),ylim=c(0,425),  xlab="Age",labels =T) 
+ggplot(players, aes(x = age)) +
+  geom_histogram(binwidth = 1, fill = "lightgray", color = "black", alpha = 1) +
+  labs(title = "Total Players by Age", x = "Age", y = "Frequency")
 ```
@@
 
-_Include citations in your Rmd file using bibtex, the list of references will automatically be placed at the end_
-
-This paper [@leek2015] discusses types of analyses.
-
-These papers [@mckay2020; @mckay2020a] are good examples of papers published using a fully reproducible setup similar to the one shown in this template.
-
-Note that this cited reference will show up at the end of the document, the reference formatting is determined by the CSL file specified in the YAML header.
-Many more style files for almost any journal [are available](https://www.zotero.org/styles).
-You also specify the location of your bibtex reference file in the YAML.
-You can call your reference file anything you like.
-
 {{< pagebreak >}}

### 2024-07-29 products/manuscript/Manuscript.qmd
Final project updates

@@
 
-The goal of this project will be to find variables related to injuries observed in the NFL.
+The goal of this project will be to conduct a descriptive analysis on variables related to injuries observed in the NFL.
 That is, find the common play types, game situation (time remaining, down, yards to go, etc.), and injured player metrics (position, height, body weight, etc.) that has a correlation to causing the resulting injury.
@@
 The primary idea behind the analysis for this project is to look at various parameters when an injury occurred and compare them to the rest of the NFL.
-If the parameters are related to the play itself, the parameter in the injury data set will be compared to cumulative play data set.
-Likewise, if the parameter is related to the injured player themselves, this parameter will be compared to the cumulative player data set.
-If there is a statistical difference between parameters, we can likely draw some conclusion about a parameter causing injuries.
+If the parameters are related to the play itself, the parameter in the "injury player data" set will be compared to "cumulative play data" set.
+Likewise, if the parameter is related to the injured player themselves, this parameter will be compared to the "cumulative player data" set.
 
 With this being said, it is imperative to ensure parameters from both data sets are in the same format in order to compare them.
-Other data cleaning items will include vetting the data sets for unexpected/missing values and converting variables to another variable that easier to use.
-For example, the play descriptions in the injury data set and cumulative play data set are very extensive in what happened during the play.
+Other data cleaning items will include vetting the data sets for unexpected/missing values and converting variables to another variable that is easier to use.
+For example, the play descriptions in the "injury player data" set and "cumulative play data" set are very extensive in what happened during the play.
 Therefore, this will be boiled down to be either a run play, short pass, etc.
@@
 ## Data aquisition
+
 The "Injury Player Data" set was found from a Github repository.
@@
 
-Lastly, the "player demographic data" set was also taken from a Github repository. Similar to the other repository, this one belongs to a group who wanted to analyze injuries in the NFL.
+Lastly, the "player demographic data" set was also taken from a Github repository. Similar to the other repository, this one belongs to a different group who wanted to analyze injuries in the NFL.
 
@@
 ## Data import and cleaning
-In both the "Injury Player Data" set and "cumulative play data" set, the play descriptions will need to be boiled down to simplified play types (pass, run, fumble, etc.).
+
+In both the "Injury Player Data" set and "cumulative play data" set, the play descriptions will need to be simplified to play types (pass, run, fumble, etc.).
 Next for the "Injury Player Data" set, the "injury_area" parameter will be summarized as upper and lower body injuries.
 This is due to the overwhelming number of injuries being classified as one or the other already so this will classify the remaining injuries as such.
-For the injury data set and the "player demographic data", player ages will need to be calculated from their birth-dates.
+For the "injury player data" set and the "player demographic data", player ages will need to be calculated from their birth-dates.
+Lastly, a new variable was created to tell whether the injured player was on the home or away team.
 
@@
 ## Statistical analysis
-First, I will select variables that show a potential difference between the injury data set and cumulative data sets through exploratory analysis.
-From these variables, I will run inferential tests for the numeric variables and a Chi-square test for the categorical variables.
+I will depict specific variables that can potentially show a relationship to number of injuries. As applicable, the same variables will be depicted from the supporting data sets for comparison.
 
@@
 ### Structures of each data set
+
+Below are glimpses into each of the three data sets showing the variables, variable types, and the first several observations as an exmaple of what the data looks like.
 
@@
 
-### Exploratory Analysis of parameters only in the *Injury* Data Set
+### Parameters only in the *Injury* Data Set
 
-Looking at the number of injuries each week in the regular season, there's a sharp drop after the first week, steadily climbs nearing the end of the season, and another drop at the end of the season.
-Intuitively, this makes sense in that players may not be physically readily for the intensity of a regular-season game despite training camps and preseason games.
+Looking at the number of injuries each week in the regular season (17 weeks), there's a sharp drop after the first week, steadily climbs nearing the end of the season, and another drop at the end of the season.
+Intuitively, this makes sense in that players may not be physically ready for the intensity of the first regular-season game despite training camps and preseason games.
 After that, the slight rise in the second half of the season could be due to exhaustion through a long season and/or teams fighting more for playoff spots.
 Lastly, the steep drop in the last week is likely due to teams knowing whether or not they have made playoffs and therefore, do not want to incur an injury to go into either the post-season or off-season.
+
 ```{r}
-hist(injuries$week.x, breaks=21,main="Injuries by Week in Regular Season", xlim=c(1,17),ylim=c(0,200),xlab="Week",labels =T)
+hist(injuries$week.x, breaks=25,main="Injuries by Week in Regular Season", xlim=c(1,22),ylim=c(0,200),xlab="Week",labels =T)
 ```
@@
 I was not expecting much but it is interesting to see about a triple in number of injuries from the lowest team (LA Rams) and the highest (LA Chargers).
+Although this graph filters out the post-season games, the four teams that made division championship games are relatively spread out across this graph and consist of the Chiefs (KC), Titans (TEN), 49ers (SF), and Packers (GB).
+Point being the most competitive teams are relatively unaffected by number of injuries.
 
@@
 
-### Exploratory Analysis of parameters in the *Injury* Data Set and *Play-by-play* Data Set
+Below depicts the number of injuries for the players that belonged to the home or away team.
+
+```{r}
+HA_freq=injuries %>% group_by(InjPlayerHomeOrAway) %>% summarize(count=n()) #filtered to only contain regular season games (weeks 1-17)
+ggplot(data = HA_freq, aes(x = reorder(InjPlayerHomeOrAway, -count), y = count))+
+  geom_col() +
+  labs(x = "Team", y="# of Injuries") + 
+  theme(axis.text.x = element_text(angle=45, hjust = 1))
+```
+
+At initial glance, the bar plot above makes it appear the number of injuries are roughly equal between the home and away teams.
+Below, I created a table to see the number of observations along with their respective proportions.
+Sure enough, there does not appear to be a large difference in the number of injuries on a team whether they are playing at home or away.
+
+```{r}
+counts = table(injuries$InjPlayerHomeOrAway) #Counts of each factor 
+proportions = prop.table(counts) #Proportions of each factor
+result = cbind(Counts = counts, Percentage = round(proportions*100,1)) #Combine results
+print(result)
+```
+
+### Parameters in the *Injury* Data Set and *Play-by-play* Data Set
 
@@
 
-There appears to be a very similar trend among the number of plays per down and the number of injuries ocurring during each down. That is, they both decrease at similar rates.
+There appears to be a very similar trend among the number of plays per down and the number of injuries occurring during each down. That is, they both decrease at similar rates.
 
@@
 
-Below are two histograms depicting first the number of injuries and then second the total # of plays by different play types.
-After a glance, there does not appear to be a significant difference in number of injuries and total # of plays ran by play type.
+Below are two histograms depicting first the number of injuries and then second the total number of plays by different play types.
+After a glance, there does not appear to be a significant difference in number of injuries and total number of plays ran by play type.
 
@@
 
-### Exploratory Analysis of parameters in the *Injury* Data Set and *Cumulative Player* Data Set
+Above, we can tell the order of play types are the same between injuries and total number of plays (e.g.
+there are more run plays causing injuries and in total relative to deep passes).
+Next, I wanted to compare the proportions of these play types.
+Below is a table of percentages of injuries and total plays by play type.
+From this, we can tell the percentages are relatively close to one another.
+It appears run plays have injuries about 4.5% less often compared to how often they're ran.
+Similarly, pass plays have injuries occur 2.5-4% more often compared to how often they are ran.
+
+```{r}
+#Injury Data set
+inj_counts = table(injuries$play_type) #Counts of each play type
+inj_proportions = prop.table(inj_counts) #Proportions of each play type
+
+#PBP data set
+total_play_counts = table(pbp$play_type) #Counts of each play type
+total_play_counts = total_play_counts[-5]
+total_play_proportions = prop.table(total_play_counts) #Proportions of each play type
+
+#Combining results
+result = data.frame(round(inj_proportions*100,1), round(total_play_proportions*100,1))
+result = result[,-3] 
+colnames(result) = c("Play Type", "Injuries Percentages", "Total Plays Percentages") #Renaming columns
+result=result[order(result$`Injuries Percentages`, decreasing = TRUE),]
+print(result)
+```
+
+Now that the play types are known, I am interested if there is a difference in plays ran between the teams with the lowest and highest number of injuries (Rams and Chargers, respectively).
+For one team having triple the number of injuries, the percentage of play types are closer than I would have expected.
+The biggest difference is between the number of deep passes.
+According to the table above, deep passes are one of the more dangerous play types relative to how often they are ran.
+Still with only a ~3% difference in deep pass plays between the two teams, there must be more factors contributing to the Chargers having so many more injuries.
+
+```{r}
+#Rams 
+ram_df = pbp[pbp$OffenseTeam == 'LA',] #Filter data set
+rams_inj_counts = table(ram_df$play_type) #Counts of each play type 
+rams_inj_proportions = prop.table(rams_inj_counts) #Proportions of each play type 
+rams_inj_proportions = rams_inj_proportions[-5]
+
+#Chargers
+cha_df = pbp[pbp$OffenseTeam == 'LAC',] #Filter data set
+cha_inj_counts = table(cha_df$play_type) #Counts of each play type 
+cha_inj_proportions = prop.table(cha_inj_counts) #Proportions of each play type
+
+#Combining results
+result = data.frame(round(rams_inj_proportions*100,1), round(cha_inj_proportions*100,1))
+print(result)
+result = result[,-3] 
+colnames(result) = c("Play Type", "Rams Percentages", "Chargers Percentages") #Renaming columns
+result=result[order(result$`Rams Percentages`, decreasing = TRUE),]
+print(result)
+```
+
+### Parameters in the *Injury* Data Set and *Cumulative Player* Data Set
 
@@
 Both histograms appear to be normally distributed.
-One interesting difference is the injured players appear to have more players just below and above the center of the histogram.
-That is, it would almost suggest players specifically under and above the mean height have a higher inclination for injury.
+One interesting difference is the injured players appear to have more players just above the center of the histogram.
+That is, it would  suggest players just above the mean height (specifically with a height of 76-77 inches) have a higher inclination for injury.
 
@@
 
-Below are the histograms depicting the players height. They are both skewed to the right with one interesting finding of a peak of injured players in the ~300lb range.
+Below are the histograms depicting the players height. They are both skewed to the right with one interesting finding of a peak of injured players in the ~310lb range.
 
@@
 
-Both distributions of players age are below and have a similar skewness to the right.
+Both distributions of players age are below and have a similar distribution.
 
@@
 
-The primary finding from the player positions is that it appears the linemen (defensive and ofensive) are injured more often relative to how many there are in the league.
+The primary finding from the player positions is that it appears the linemen (defensive and offensive) are injured more often relative to how many there are in the league.
 
@@
 
-## Basic statistical analysis
-
-From the exploratory analysis, it appears there three numeric variables that may show a difference and one categorical.
-The three numeric variables are quarter, height, and weight while the single categorical variable is player position.
+To make findings about the positions more clear, a table of percentages for each position can be found below for injured players and all of players.
+From the graphs, we could easily tell the defensive linemen and offensive linemen have higher percentages of being injured relative to their proportions in the league.
+The table below suggests linebackers, wide receivers, and running backs are injured 4-7% less relative to their respective proportions in the league.
 
 ```{r}
-#Inferential test looking at Quarter
-shapiro.test(injuries$qtr) #p-value<0.05. Reject Ho & don't assume normal.
-shapiro.test(sample(pbp$Quarter, size=5000, replace=F)) #p-value<0.05. Reject Ho & don't assume normal.
-#Will do Wilcoxon rank-sum test b/c of non-normal data
+#Injury Data set
+inj_counts = table(injuries$position) #Counts of each position
+inj_proportions = prop.table(inj_counts) #Proportions of each position
 
-wilcox.test(injuries$qtr, pbp$Quarter, exact=F) #p-value<0.05. Reject Ho and therefore, one group tends to have larger values than the other
+#Players Data set
+total_play_counts = table(players$position) #Counts of each position
+total_play_proportions = prop.table(total_play_counts) #Proportions of each position
 
-```
-
-```{r}
-#Inferential test looking at Height
-shapiro.test(injuries$height) #p-value<0.05. Reject Ho & don't assume normal.
-shapiro.test(players$height) #p-value<0.05. Reject Ho & don't assume normal.
-#Will do Wilcoxon rank-sum test b/c of non-normal data
-
-wilcox.test(injuries$height, players$height, exact=F) #p-value<0.05. Reject Ho and therefore, one group tends to have larger values than the other
-
-```
-
-```{r}
-#Inferential test looking at Weight
-shapiro.test(injuries$weight) #p-value<0.05. Reject Ho & don't assume normal.
-shapiro.test(players$weight) #p-value<0.05. Reject Ho & don't assume normal.
-#Will do Wilcoxon rank-sum test b/c of non-normal data
-
-wilcox.test(injuries$weight, players$weight, exact=F) #p-value<0.05. Reject Ho and therefore, one group tends to have larger values than the other
-
-```
-```{r}
-#Chi-square test looking at player position
-factor_levels = levels(injuries$position)
-chisq.test(sample(injuries$position, size=1303, replace=F), players$position) #p-value>0.05. Accept Ho and therefore, the two data sets have equal proportions among the player positions
-
-```
-
-## Full analysis
-
-_Use one or several suitable statistical/machine learning methods to analyze your data and to produce meaningful figures, tables, etc.
-This might again be code that is best placed in one or several separate R scripts that need to be well documented.
-You want the code to produce figures and data ready for display as tables, and save those.
-Then you load them here._
-
-Example @tbl-resulttable2 shows a summary of a linear model fit.
-
-```{r}
-#| label: tbl-resulttable2
-#| tbl-cap: "Linear model fit table."
-#| echo: FALSE
-resulttable2 = readRDS(here("results","tables","resulttable2.rds"))
-knitr::kable(resulttable2)
+#Combine results
+result = data.frame(round(inj_proportions*100,1), round(total_play_proportions*100,1))
+result = result[,-3] 
+colnames(result) = c("Player Position", "Injuries Percentages", "Total Percentages") #Renaming columns
+result=result[order(result$`Injuries Percentages`, decreasing = TRUE),]
+print(result)
 ```
@@
 ## Summary and Interpretation
-_Summarize what you did, what you found and what it means._
+
+The purpose of this analysis was to find the variables within these data sets that have a correlation to causing injuries.
+First, what kind of injuries?
+Although we have limited information on the specifics, it should be noted lower body injuries occur more often than upper body and head injuries.
+
+The parameters that appeared to have an effect on injuries would consist of the week in the season, quarter, play type, and player position.
+The finding from each parameter is summarized below.
+It is worth mentioning players height and weight also had an effect but would be highly correlated to the player position.
+1.
+Week during season - A spike of injuries at the beginning of the season followed by a steady increase throughout the rest of the season.
+Likely due to being unprepared for a regular season game and then more injuries throughout the season as fatigue sets in.
+2. Quarter of a game - Injuries steadily increases throughout the course of a game likely due to fatigue. 
+3. Play type - There is a 2-5% difference in number of injuries compared to the total number ran. Runs have slight lower chance of injury while pass plays have a slight higher chance. 
+4. Player Position - By and large, linemen (offensive and defensive) have a greater chance of injury.
+
+A notable mention would be number of injuries per team during the regular season.
+One can expect a difference due to coaching staff, team culture, ambient conditions of home field but the LA Chargers having three times the number of injuries as the LA Rams was unexpected.
 
 ## Strengths and Limitations
-_Discuss what you perceive as strengths and limitations of your analysis._
+
+Strengths of this analysis are that it was thorough as a descriptive analysis.
+As applicable, parameters from the injury data set were compared to the population of the NFL to see if there were any meaningful differences.
+
+With this being said, this analysis consisted of comparing plots and percentages.
+The next step of this would be to conduct inferential analysis to solidify the findings in this initial analysis.
+Furthermore, this analysis would be interesting to conduct on multiple years worth of data.
+This analysis was of a single NFL season and therefore, it would be interesting to see a trend in this data across different seasons.
 
 ## Conclusions
-_What are the main take-home messages?_
+
+First, the increased number of injuries being in the lower body is a reminder to coaches and players to stay hydrated and stretch before games to avoid pulled muscles.
+Although the data does not call out pulled muscles being a major contributor, this is a good reminder to limit injuries if you can.
+
+As for coaches, be aware of the variables discussed in this analysis that have a higher chance of injuries.
+Injuries occur more often during the first week of the season so perhaps adjustments can be made in the preseason to better prepare the players.
+The coaches with a higher proportion of injuries should consider collaborating with the coaches of lower numbers to see if there is something different they can do to prevent these for the players.
+Preseason rituals are likely proprietary but I would think it would be in the interest of the NFL to investigate this and come up with a best practices for the teams.
 

### 2024-07-29 code/processing-code/processingcode.R

@@ -54,6 +54,7 @@ dplyr::glimpse(injuries)
   #Summarize "description" into "play_type"
   #Summarize "injury_area" as upper vs lower body into new parameter
   #Create new field calculating age from birthdate (mm/dd/yyyy)
+  #Create new paramter to tell whether injured player is on home or away team 
 dplyr::glimpse(pbp) 
   #Convert "gamedate" to date (from char), 
   #remove empty columns (X, X.1,X.2,X.3), 
@@ -145,14 +146,25 @@ for (i in 1:nrow(d1)) {
 ## ---- cleaninjuriesdata3 --------
 #Create new field calculating age from birthdate (mm/dd/yyyy).
 is.na(d1$birth_date) = NA #Some NA values were not recognized by R as NA or anything else 
-d1$birth_date <- as.Date(d1$birth_date, format = "%m/%d/%Y", tryFormats = c("%m/%d/%Y")) 
-d1$game_date <- d1$game_date+as.Date("1899-12-30", "%Y-%m-%d") #Convert game_date from integer of # days after 12/30/1899 to date
+d1$birth_date = as.Date(d1$birth_date, format = "%m/%d/%Y", tryFormats = c("%m/%d/%Y")) 
+d1$game_date = d1$game_date+as.Date("1899-12-30", "%Y-%m-%d") #Convert game_date from integer of # days after 12/30/1899 to date
 d1$age = NA
 d1$age = round(as.numeric(difftime(d1$game_date, d1$birth_date, units = 'weeks'))/52, digits=1)
 
 ## ---- cleaninjuriessdata4 --------
 #Character parameters may also need to be converted to factor. 
-d1$position <- as.factor(d1$position)
+d1$position = as.factor(d1$position)
+
+## ---- cleaninjuriessdata5 --------
+#New variable to tell whether injured player is on home team or not 
+d1$InjPlayerHomeOrAway = NA
+for (i in 1:nrow(d1)) {
+  if(is.na(d1$team[i]) || is.na(d1$home_team[i]) || is.na(d1$away_team[i])) {d1$InjPlayerHomeOrAway[i] = NA}
+  else if(d1$team[i] == d1$home_team[i]) {d1$InjPlayerHomeOrAway[i] = 'Home'}
+  else if(d1$team[i] == d1$away_team[i]) {d1$InjPlayerHomeOrAway[i] = 'Away'}
[…]
